---
title: "Demonstrate Statistical Learning and Data Science by R Markdown"
author: "Xinwei Deng (xdeng@vt.edu), Virginia Tech"
date: "`r Sys.Date()`"
header-includes: 
   - \usepackage{amsmath,bm,bbm,amssymb,mathrsfs}
   - \usepackage{amsthm,amscd,float}             
   - \usepackage{graphicx, multirow, graphics}
   - \usepackage{comment}
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

* R Markdown is a powerful intergation of R and techchnical writing.
* It can produce a useful report for data analytics with informative elaboration.
* It is also very convenient to incorport the Latex math writing.
    - For example, assume that we like to generate the iid sample from the multivariate normal distribution $X \sim N(\mu, \Sigma)$. 
    - We can also call the latex package to create equation, table and figures.

```{r, eval=TRUE, echo = TRUE}
library(MASS)
n = 1000
p = 5
mu = rep(0,p)

rho = -0.5
Sigma_mat = diag(rep(1,p));
for( j in 1: (p-1) )
{
	Sigma_mat[j,j+1] = rho;
	Sigma_mat[j+1,j] = rho;	
}

Data = mvrnorm(n, mu, Sigma_mat);
plot(Data[,1],Data[,2], xlab = "x1", ylab = "y")
```

From the above plot, one can see the shape of the normal distribution. 

## Model Estimation 

* The R markdown can also be used to condut the data analysis and show the model estimation results. 
* For example, one can consider the linear model as
    \begin{align}
    Y = \beta_{0} + \beta_{1} X+\epsilon,
    \end{align}
where $\epsilon$ is the error terms following the zero-mean normal distribution $N(0,\sigma^{2})$.
```{r, eval=TRUE, echo = TRUE}
y = Data[,2]
x1 = Data[,1]
x2 = Data[,3]
x3 = Data[,4]
x4 = Data[,5]

mod1 = lm(y~x1);
mod1
summary(mod1);
```
It is seen that the estimated slope $\hat{\beta}$ is `r mod1$coef[2]`, which is very close to $-0.5$.

The linear model is called linear in terms of $\beta$.
    \begin{align}
    Y = \beta_{0} + \sum_{j-=1}^{p} \beta_{j} x_{j} +\epsilon,
    \end{align}
    
```{r, eval=TRUE, echo = TRUE}
mod2 = lm(y~x1+x2+x3+x4);
summary(mod2);
```
Based on the course material, it is known that
    \begin{align}
    E(y|\mathbf{x}) = \mathbf{x}'\Sigma_{xx}^{-1}\Sigma_{xy} = -\mathbf{x}' \frac{\Omega_{xy}}{\Omega_{yy}} .
    \end{align}
where $\Omega$ is the inverse of the covaraince matrix $\Sigma$.
```{r, eval=TRUE, echo = TRUE}
XY_data = cbind(x1,x2,x3,x4,y);

Omega_mat = solve(cov(XY_data))
Omega_xy = Omega_mat[1:4,5]
Omega_yy = Omega_mat[5,5]

beta_hat = -Omega_xy/Omega_yy
print(beta_hat)
```
The above analysis shows that the values of estimated coefficient $\hat{\beta}$ is the same as the second row in the inverse covariance matrix of $\Sigma$.

## Data and Model Visualization

* The R Shiny can also be incorporated into the R markdown to dynamically elaborate the data analysis. 
For example, we can run the following code.
```{r}
library(shiny)

# Define UI, the user interface ----
ui <- fluidPage(
  ## Put a title here:
  titlePanel("Data (Teach.csv) Table"),
  
  sidebarLayout(
    
    sidebarPanel(
      
      ## Place the widget here:
      numericInput(inputId = "num", label = "number of observations to display:",
                   value = 6, min = 0, max = 90, step = 1)
    ),
    
    ## Specify the table output here:
    mainPanel(tableOutput(outputId = "df"))
    )
  )

# Define server logic ----
server <- function(input, output)
  {
  
  ## Grab the data
  df <- read.table("teach.csv",
                   sep = ",", header = TRUE)
  df$degree <- as.factor(df$degree)
  
  ## Generate the dataset with user specified number of rows.
  ## Notice it is in a reactive() function. 
  dat <- reactive({
    return(df[1:input$num, ])
  }) 
  
  ## Create the output of a table:
  output$df <- renderTable(dat())
  }

# Run the app ----
shinyApp(ui = ui, server = server)
```
We can see that for the R Shiny there are two key components:

* The ${\bf ui}$ component: it plays the functionality of structure design
* The ${\bf server}$ component: it decides what to be conducted in the `ui` function.

Now we can show another example for elaborating the model fitting.

```{r}
library(shiny)
library(ggplot2)
library(MASS)
options(stringsAsFactors = FALSE)

# Define UI ----
ui <- fluidPage(
  titlePanel("Teach.csv Example"),
  sidebarLayout(
    sidebarPanel(
        checkboxInput(inputId = "smooth", label = "Fit Regression"),
        ## Only until smooth is checkboxed, will the method dropdown appear
        ## A conditional panel is a 'widget' with a condition specified.
        conditionalPanel(
            condition = "input.smooth == true",
            selectInput(inputId = "smoothMethod", label = "Method",
                        list("lm", "glm", "gam", "loess", "rlm"))
            )
        ),
    mainPanel(plotOutput(outputId = "plot"))))

# Define server logic ----
server <- function(input, output){
  df <- read.table("teach.csv",
                   sep = ",", header = TRUE)
  
  # Using ggplot to draw the scatter plot and regression line.
  ## One plot puts a regression line, and other is just a scatterplot.
  output$plot <- renderPlot({
      if(input$smooth){
          ggplot(data = df) +
              geom_jitter(mapping = aes(x = months, y = salary)) +
              # geom_smooth(mapping = aes(x = months, y = salary)) +
              geom_smooth(mapping = aes(x = months, y = salary),
                          method = input$smoothMethod) +
              xlab("Month") +
              ylab("Salary") +
              ggtitle("Salary Versus Months")
      } else{
          ggplot(data = df) +
              geom_jitter(mapping = aes(x = months, y = salary)) +
              xlab("Month") +
              ylab("Salary") +
              ggtitle("Salary Versus Months")
      }
  })
}
# Run the app ----
shinyApp(ui = ui, server = server)
```
One can also easily obtain the residual plot to evaluate the model assumptions. For example, we consider the linear model with $Y$ to be salary, and the other variables to be predictor varaibles $x_{1}, \ldots, x_{p}$.

```{r}
library(broom)
library(ggplot2)
# fit the data
Tdata <- read.table("teach.csv", sep = ",", header = TRUE)
mod3 = lm(salary~months+sex+marry+degree+type+train+brk, data = Tdata)
summary(mod3)

# Define UI ----
ui <- fluidPage(
  
  titlePanel("Residual Analysis"),
  sidebarLayout(
    
    sidebarPanel(),
      
    mainPanel = (
      tabsetPanel(
        # tabset for different plots
        tabPanel("Residual Plot", plotOutput(outputId = "residplot")), 
        tabPanel("Histogram", plotOutput(outputId = "histplot")),
        tabPanel("QQ Plot", plotOutput(outputId = "qqplot"))
      )
    ),
    )
  
 )

# Define server logic ----
server <- function(input, output) {
  
  # residual plot
  output$residplot <- renderPlot({
    df <- augment(mod3)
    ggplot(data = df) + 
      geom_jitter(mapping = aes(x = .fitted, y = .resid)) +
      geom_abline(intercept = 0, slope = 0) +
      ggtitle("Residuals Versus Fitted") +
      xlab("Fitted") +
      ylab("Residuals")

  })
  # histogram of residuals
  output$histplot <- renderPlot({
    df <- augment(mod3)
    ggplot(data = df) + 
      geom_histogram(mapping = aes(x = .resid), bins = 50) +
      ggtitle("Histogram of Residuals") +
      xlab("Residuals")
  })
  # qqplot for normality check
  output$qqplot <- renderPlot({
    df <- augment(mod3)
    ggplot(data = df) +
      stat_qq(mapping = aes(sample = .resid)) +
      ggtitle("Normal QQ Plot") +
      ylab("Standardized Residuals") +
      xlab("Theoretical Quantiles")
  })
  
}
# Run the app ----
shinyApp(ui = ui, server = server)
```
In a short summary, the R markdown is useful for technical writing, with a good integration of data analytics and data visiualization. 




    



